Fitting metabolic models to dissolved oxygen data: The Estuarine BAyesian Single-station Estimation method (EBASE)

### 2024-03-21 manu-draft.qmd
SE edits

$$
Z\frac{dC}{dt} = P - R - D
$$ {#eq-ebase}

where the terms on the right side of the equation are gross primary production ($P$), ecosystem respiration ($R$), and the net upward diffusive gas flux at the air--water interface ($D$), respectively, processes that result in the change in dissolved oxygen concentration ($dC$) per unit time ($dt$). Both $P$ and $R$ are expressed as positive values, whereas $D$ can be positive (outgassing) or negative (ingassing). The equation has dimensions of moles per unit area and time (e.g., $\text{mmol}~\text{m}^{-2}~\text{d}^{-1}$). $P$ is modeled as $aPAR$, where $PAR$ is the surface photosynthetically active radiation and $a$ is the light efficiency; $R$ is constant; and $D$ is modeled in the standard manner [e.g, @sarmiento2013] as $k_w\left(C_{sat} - C\right)$, where $k_w$ is the gas transfer velocity and $C_{sat}$ is the saturation (or equilibrium) concentration, calculated from salinity and temperature [@Garcia92]. The gas transfer velocity is modeled using the formulation in @Wanninkhof14:

@@

The resulting synthetic time series of light efficiency, dissolved oxygen budget terms ($P$, $D$, and $R$), and dissolved oxygen concentration are shown in @fig-synapa. The light efficiency has a complex structure, with maxima in the summer, fall, and winter, and a strong minimum in spring (Figure [-@fig-synapa]a). The oxygen budget terms (Figure [-@fig-synapa]b) reveal that (1) photosynthesis has a structure similar to, yet noisier than, that of light efficiency, reflecting the synoptic-scale variability of $PAR$; (2) respiration has the expected structure of being greatest during the summer [@caffrey2004], stimulated by greater autochthonous organic matter availability and high temperature; and (3) the air--water gas flux mimic productivity on short time scales, with ingassing during high productivity and outgassing during low productivity. The synthetic dissolved oxygen time series shows high variability at multiple time scales (Figure [-@fig-synapa]c). There is a prominent diurnal cycle that is superimposed on subseasonal variability, presumably driven by radiation and winds.

>>>>

$$
Z\frac{dC}{dt} = P - R + D
$$ {#eq-ebase}

where the terms on the right side of the equation are gross primary production ($P$), ecosystem respiration ($R$), and the net upward diffusive gas flux at the air--water interface ($D$), respectively, processes that result in the change in dissolved oxygen concentration ($dC$) per unit time ($dt$). Both $P$ and $R$ are expressed as positive values, whereas $D$ can be positive (ingassing) or negative (outgassing). The equation has dimensions of moles per unit area and time (e.g., $\text{mmol}~\text{m}^{-2}~\text{d}^{-1}$). $P$ is modeled as $aPAR$, where $PAR$ is the surface photosynthetically active radiation and $a$ is the light efficiency; $R$ is constant; and $D$ is modeled in the standard manner [e.g, @sarmiento2013] as $k_w\left(C_{sat} - C\right)$, where $k_w$ is the gas transfer velocity and $C_{sat}$ is the saturation (or equilibrium) concentration, calculated from salinity and temperature [@Garcia92]. The gas transfer velocity is modeled using the formulation in @Wanninkhof14:

@@

The resulting synthetic time series of light efficiency, dissolved oxygen budget terms ($P$, $D$, and $R$), and dissolved oxygen concentration are shown in @fig-synapa. The light efficiency has a complex structure, with maxima in the summer, fall, and winter, and a strong minimum in spring (Figure [-@fig-synapa]a). The oxygen budget terms (Figure [-@fig-synapa]b) reveal that (1) photosynthesis has a structure similar to, yet noisier than, that of light efficiency, reflecting the synoptic-scale variability of $PAR$; (2) respiration has the expected structure of being greatest during the summer [@caffrey2004], stimulated by greater autochthonous organic matter availability and high temperature; and (3) the air--water gas flux mimic productivity on short time scales, with outgassing during high productivity and ingassing during low productivity. The synthetic dissolved oxygen time series shows high variability at multiple time scales (Figure [-@fig-synapa]c). There is a prominent diurnal cycle that is superimposed on subseasonal variability, presumably driven by radiation and winds.
